---
title: "Lab 4"
author: "My Name"
output:
  pdf_document:
    fig_caption: yes
    number_sections: true
header-includes:
  - |
    ```{=latex}
    \usepackage{fvextra}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{
      breaksymbolleft={}, 
      showspaces = false,
      showtabs = false,
      breaklines,
      commandchars=\\\{\}
    }
    ```
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy.opts=list(width.cutoff=70),
  tidy=TRUE
)
```

# Lab 4 (Set 2): Distributions {-}

# Prepare a corpus

## Load the needed packages

```{r setup, message = FALSE, error=FALSE, warning=FALSE}
library(tidyverse)
library(quanteda)
library(quanteda.textstats)
```

Load data:

```{r}
load("../data/sample_corpus.rda")
load("../data/multiword_expressions.rda")
```

Load functions:

```{r}
source("../R/dispersion_functions.R")
```


## Load a corpus

The **cmu.textstat** package comes with some data sets including a data table with a column of document ids and a column of texts. Such a table is easy to create from text data on your own local drive.

To do so, you would organize plain .txt files into a directory and use the **readtext( )** function from the **readtext** package.

```{r make_data}
sc_df <- sample_corpus
```

To peek at the data, we'll look at the first 100 characters in the "text" column of the first row:

```{r echo=FALSE}
knitr::kable(substr(sc_df$text[1], start = 1, stop = 100), col.names = "")
```

Make a corpus object.
```{r create_corpus}
sc <- corpus(sc_df)
```

And check the result:

```{r echo=FALSE}
knitr::kable(head(sc %>% summary()), caption = "Partial summary of sample corpus.")
```

## Document variables (Name your files systematically!)

Note the names of the text files encode important meta-data: in this case, the names of text types similar to the Corpus of Contemporary American English.

This is **extremely important**. When you build your own corpora, you want to purposefully and systematically name your files and organize your directories. This will save you time and effort later in your analysis.

We are now going to extract the meta-data from the file names and pass them as a variable.

```{r extract_variable}
doc_categories <- str_extract(sc_df$doc_id, "^[a-z]+")
```

Check the result:

```{r echo=FALSE}
knitr::kable(doc_categories %>% unique(), col.names = "", caption = "Document categories.")
```

We will now assign the variable to the corpus. The following command might look backwards, with the function on the left hand side of the **<-** operator. That is because it's an accessor function, which lets us add or modify data in an object. You can tell when a function is an accessor function like this because its help file will show that you can use it with **<-**, for example in **?docvars**.

```{r}
docvars(sc, field = "text_type") <- doc_categories
```

And check the summary again:

```{r echo=FALSE}
knitr::kable(head(sc %>% summary()), caption = "Partial summary of sample corpus.")
```

## Tokenize the corpus

We'll use **quanteda** to tokenize. And after tokenization, we'll convert them to lower case. *Why do that here?* As a next step, we'll being combining tokens like *a* and *lot* into single units. And we'll be using a list of expressions that isn't case sensitive.

```{r tokenize}
sc_tokens <- tokens(sc, include_docvars=TRUE, remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE, what = "word")

sc_tokens <- tokens_tolower(sc_tokens)
```

## Multi-word Expressions

An issue that we run into frequently with corpus analysis is what to do with multi-word expressions. For example, consider a common English quantifier: "a lot". Typical tokenization rules will split this into two tokens: *a* and *lot*. But counting *a lot* as a single unit might be important depending on our task. We have a way of telling **quanteda** to account for these tokens.

All that we need is a list of multi-word expressions.

The **cmu.textstat** comes with an example of an mwe list called **multiword_expressions**:

```{r echo=FALSE}
knitr::kable(tail(multiword_expressions), col.names = "", caption = "Examples of multi-word expressions.")
```

The **tokens_compound()** function looks for token sequences that match our list
and combines them using an underscore.

```{r compound_tokens}
sc_tokens <- tokens_compound(sc_tokens, pattern = phrase(multiword_expressions))
```

## Create a document-feature matrix

With our tokens object we can now create a document-feature-matrix using the
**dfm()** function. As a reminder, a **dfm** is table with one row per document in the corpus, and one column per unique token in the corpus. Each cell contains a count of how many times a token shows up in that document.

```{r dfm}
sc_dfm <- dfm(sc_tokens)
```

Next we'll create a **dfm** with proportionally weighted counts.

```{r}
prop_dfm <- dfm_weight(sc_dfm, scheme = "prop")
```

# Token distributions

## Distributions of *the*

Let's start by selecting frequencies of the most common token in the corpus:

```{r}
freq_df <- textstat_frequency(sc_dfm) %>%
  data.frame(stringsAsFactors = F)
```

```{r echo=FALSE}
knitr::kable(freq_df[1:10,], caption = "The 10 most frequent tokens in the sample corpus.")
```

From the weighted **dfm**, we can select any token that we'd like to look at more closely. In this case, we'll select the most frequent token: *the*.

After selecting the variable, we will convert the data into a more friendly data structure.

There are easier ways of doing this, but the first bit of the code-chunk allows us to filter by rank and return a character vector that we can pass. This way, we can find a word of any arbitrary rank.

Also note how the **rename()** function is set up. Let's say our token is *the*. The **dfm_select()** function would result with a column named **the** that we'd want to rename **RF**. So our typical syntax would be: **rename(RF = the)**. In the chunk below, however, our column name is the variable **word**. To pass that variable to **rename**, we use **!!name(word)**.

```{r select, warning=FALSE, message=FALSE}
word <- freq_df %>% 
  filter(rank == 1) %>% 
  dplyr::select(feature) %>%
  as.character()

word_df <- dfm_select(prop_dfm, word, valuetype = "fixed") # select the token

word_df <- word_df %>% 
  convert(to = "data.frame") %>% 
  cbind(docvars(word_df)) %>% 
  rename(RF = !!as.name(word)) %>% 
  mutate(RF = RF*1000000)
```

With that data it is a simple matter to generate basic summary statistics using the **group_by()** function:

```{r}
summary_table <- word_df %>% 
  group_by(text_type) %>%
  summarize(MEAN = mean(RF),
              SD = sd(RF),
              N = n())
```

```{r echo=FALSE}
knitr::kable(summary_table, digits = 2, caption = "Means and standard deviations by text-type.")
```

And we can inspect a histogram of the frequencies. To set the width of our bins we'll use the Freedman-Diaconis rule. The bin-width is set to:
$$h = 2 x \frac{IQR(x)}{n^{1/3}}$$

So the number of bins is (max-min)/h, where n is the number of observations, max is the maximum value and min is the minimum value.

```{r bin_width}
bin_width <- function(x){
  2 * IQR(x) / length(x)^(1/3)
  }
```

Now we can plot a histogram. We're also adding a dashed line showing the mean. Note we're also going to use the **scales** package to remove scientific notation from our tick labels.

```{r the_histogram, fig.height=2.5, fig.width=7, fig.cap="Histogram of the token \\textit{the}."}
ggplot(word_df,aes(RF)) + 
  geom_histogram(binwidth = bin_width(word_df$RF), colour="black", fill="white", size=.25) +
  geom_vline(aes(xintercept=mean(RF)), color="red", linetype="dashed", size=.5) +
  theme_classic() +
  scale_x_continuous(labels = scales::comma) +
  xlab("RF (per mil. words)")
```

## Distributions of *the* and *of*

Now let's try plotting histograms of two tokens on the same plot. First we're going to use regular expressions to select the columns. The carat or hat **^** looks for the start of line. Without it, we would also get words like "blather". The dollar symbol **$** looks for the end of a line. The straight line | means OR. Think about how useful this flexibility can be. You could, for example, extract all words that end in *-ion*.

```{r warning=FALSE}
# Note "regex" rather than "fixed"
word_df <- dfm_select(prop_dfm, "^the$|^of$", valuetype = "regex")

# Now we'll convert our selection and normalize to 10000 words.
word_df <- word_df %>% 
  convert(to = "data.frame") %>%
  mutate(the = the*10000) %>%
  mutate(of = of*10000)

# Use "pivot_longer" to go from a wide format to a long one
word_df <- word_df %>% 
  pivot_longer(!doc_id, names_to = "token", values_to = "RF") %>% 
  mutate(token = factor(token))
```

Now let's make a new histogram. Here we assign the values of color and fill to the "token" column. We also make the columns a little transparent using the "alpha" setting.

```{r fig.height=2.5, fig.width=7, fig.cap="Histogram of the tokens \\textit{the} and \\textit{of}."}
ggplot(word_df,aes(x = RF, color = token, fill = token)) + 
  geom_histogram(binwidth = bin_width(word_df$RF), alpha=.5, position = "identity") +
  theme_classic() +
  xlab("RF (per mil. words)") +
  theme(axis.text = element_text(size=5))
```

If we don't want overlapping histograms, we can use **facet_wrap()** to split the plots.

```{r fig.height=2.5, fig.width=7, fig.cap="Histogram of the tokens \\textit{the} and \\textit{of}."}
ggplot(word_df,aes(x = RF, color = token, fill = token)) + 
  geom_histogram(binwidth = bin_width(word_df$RF), alpha=.5, position = "identity") +
  theme_classic() +
  theme(axis.text = element_text(size=5)) +
  theme(legend.position = "none") +
  xlab("RF (per mil. words)") +
  facet_wrap(~ token)
```

___
\begin{center}
STOP!\\
COMPLETE TASK 1
\end{center} 
___

# Dispersion

We can also calculate dispersion, and there are a variety of measures at our disposal. Our toolkit has several functions for producing these calculations.

For example, we can find the dispersion of any specific token:

```{r token_dispersions}
the <- dispersions_token(sc_dfm, "the") %>% unlist()
```

```{r echo=FALSE}
knitr::kable(the, digits = 3, caption = "Dispersion measures for the token \\textit{the}.", "simple", col.names = "")
```

And let's try another token to compare:

```{r}
data <- dispersions_token(sc_dfm, "data") %>% unlist()
```

```{r echo=FALSE}
knitr::kable(data.frame(the = the["Deviation of proportions DP"], data = c(data["Deviation of proportions DP"])), digits = 3, caption = "Deviation of Proportions for the tokens \\textit{the} and \\textit{data}.", "simple")
```

___
\begin{center}
STOP!\\
COMPLETE TASK 2
\end{center} 
___


## Dispersions for all tokens

We can also calculate selected dispersion measures for all tokens using **dispersions_all()**:

```{r all_dispersions, warning=FALSE, message=FALSE}
d <- dispersions_all(sc_dfm)
```

```{r echo=FALSE}
knitr::kable(head(d), digits = 3, caption = "Dispersion measures for all tokens.", "simple")
```

## Generating a frequency table

Alternatively, **frequency_table()** returns only Deviation of Proportions and Average Reduced Frequency.

Note that ARF requires a tokens object and takes a couple of minutes to calculate.

```{r frequency_table}
ft <- frequency_table(sc_tokens)
```

```{r echo=FALSE}
knitr::kable(head(ft), digits = 3, caption = "Frequency and dispersion measures for all tokens.", "simple")
```

___
\begin{center}
STOP!\\
COMPLETE TASK 3
\end{center} 
___


# Zipf's Law

Let's plot rank against frequency for the 100 most frequent tokens in the sample corpus.

```{r fig.height=2.5, fig.width=7, fig.cap="Token rank vs. frequency."}
ggplot(freq_df %>% filter(rank < 101), aes(x = rank, y = frequency)) +
  geom_point(shape = 1, alpha = .5) +
  theme_classic() +
  ylab("Absolute frequency") +
  xlab("Rank")
```

The relationship you're seeing between the rank of a token and it's frequency holds true for almost any corpus and is referred to as **Zipf's Law** (see Brezina pg. 44).

___
\begin{center}
STOP!\\
COMPLETE TASK 4
\end{center} 
___


