---
title: "Homework_02"
output: pdf_document
author: "My Name"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploring Two Different Genres

For this homework assignment, you'll be doing some descriptive analysis of two different corpora, which contain two very different genres. I'll give you a start with some code for generating a document feature matrix and word frequencies from a corpus of randomly sampled tweets.

First, be sure to set the **Knit Directory** as the **Project Directory** using the **Knit** menu above. Otherwise, the data links won't work.

```{r twitter, include=FALSE, error=FALSE}
# load the quanteda and tidyverse libraries
library(quanteda)
library(tidyverse)

# save our text
twitter_df <- read_csv("data/twitter_data/tweet_sample.csv")

# the twitter data has a lot of emojis and other special characters
# that are read as letters and numbers between angle brackets < >
# so were going to use str_remove_all() so we don't count them.
twitter_df <- twitter_df %>%
  mutate(text = str_remove_all(text, '<\\S+>'))

# convert the text into a corpus object
twitter_corpus <- corpus(twitter_df)

# tokenize it
twitter_tokens <- tokens(twitter_corpus, what = "word", remove_punct = TRUE, remove_url = T)

# create a document frequency matrix
twitter_dfm <- dfm(twitter_tokens)

# make a table of token counts
twitter_counts <- textstat_frequency(twitter_dfm) %>%
  mutate(freq_norm = (frequency/sum(frequency))*10000)

```

## Frequency Table

We can, as before, generate a table of frequencies. This time for the ten most frequent tokens:

```{r counts, echo=TRUE}
knitr::kable(twitter_counts[1:10, ], caption = 'A caption')
```

## Homework

1. Try creating a histogram of the token *the* from the twitter dfm. **Hint**: you can draw from the code that we used in 03_dispersions.R and note that the result may not be what you expect. Can you plot your reults?

```{r}
# code goes here
```

2. Calculate the DP norm for the token *the* and report the result.

```{r, echo=TRUE}
source("functions/dispersion_functions.R")
```

3. Compare the distributions and dispersions of *the* in the Twitter corpus to those in MICUSP. Report the results.

```{r}
# code goes here
```

4. Posit and explanation for your findings. **Hint**: Think carefully about the **function** of the determiner in English.

5. Be sure to hide/show your code chunks intentionally, fix any captions, clearly label axes, and thoughtfully design your visualizations.
